---
author: "Steve Rozen"
output: html_document
date: '`r Sys.Date()`'
params:
  bgfactor: 2
  showrecon: TRUE
  showbgspec: TRUE
  onefhigh: 2.5
  verbose: TRUE
  count.nbinom.size: 20
  def.fig.wid: 4
  
title: Background subtraction for various platinum-based chemotherapeutics  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Parameters

### Inflation for mean number of background mutations = `r params[["bgfactor"]]`.

### Dispersion parameter for number of background mutations = `r params[["count.nbinom.size"]]`.

## Introduction

This is an example of background subtraction using the mSigBG package.

In this example we subtract the background signatures of
HepG2 and MCF-10A cells from the spectra of cells treated with
three platinum-based chemotherapies.

`r ifelse(params$verbose, "## Libraries, graphics parameters, and helper functions", "")`


```{r, libraries, echo=params$verbose}
library(mSigBG)
library(ICAMS)
library(BSgenome.Hsapiens.1000genomes.hs37d5)
```

```{r, util_func, echo=params$verbose}

set.initial.par <- function(catalog) {
  par(pin = c(3*ncol(catalog), 1))
  par(mfrow = c(ncol(catalog), 1))
  par(mar = c(2, 4, 4, 2))
  par(cex = 0.8)
  par(cex.main = 1.4) 
}

last.bottom.mar <- 4.5
```


```{r, define_pcat1, echo=params$verbose}
# Plot a single catalog
pcat1 <- function(catalog, ylim = NULL) {
  par(pin = c(5, 1))
  par(mar = c(5, 4, 5, 4))
  par(cex = 0.8)
  par(cex.main = 1.4)
  bp <- ICAMS::PlotCatalog(catalog[ , 1, drop = FALSE],
                           upper   = TRUE,
                           xlabels = TRUE,
                           ylim    = ylim
                           )
  return(invisible(bp$plot.object))
}
```

```{r, define_pcat, echo=params$verbose}
pcat <- function(catalog) {
  set.initial.par(catalog)
  xlabels <- FALSE
  for (i in 1:ncol(catalog)) {
    if (FALSE && i == ncol(catalog)) { # FALSE an experiment
      xlabels <- TRUE
      par(mar = c(last.bottom.mar, 4, 3, 2))
    }
    
    ICAMS::PlotCatalog(catalog[ , i, drop = FALSE],
                       upper = (i == 1),
                       xlabels = xlabels)
  }
}
```

```{r, define_stacked_pcat, echo=params$verbose}
stacked.pcat <- function(bg.catalog, target.catalog, set.neg.zero = TRUE) {
  
  set.initial.par(bg.catalog)
  xlabels <- FALSE

  for (i in 1:ncol(bg.catalog)) {
    if (FALSE && i == ncol(bg.catalog)) { # FALSE an experiment
      xlabels <- TRUE
      par(mar = c(last.bottom.mar, 4, 3, 2))
    }
    
    target.cat.1 <- target.catalog[ , i, drop = FALSE]
    bg.cat.1     <- bg.catalog[ , i, drop = FALSE]    
    
    zeros <- which(target.cat.1 < 0)
    neg.catalog <- target.cat.1
    
    if (set.neg.zero) {
      target.cat.1[zeros] <- 0
    }
    neg.catalog[-zeros] <- 0
    neg.catalog <- neg.catalog * -1

    to.plot <- cbind(bg.cat.1, target.cat.1)
    
    bp <- ICAMS::PlotCatalog(
      to.plot,
      upper   = (i == 1),
      xlabels = xlabels)$plot.object
    
    bpz <- bp[zeros]
    # message(paste(bpz, collapse = " "))

    y0 <- bg.cat.1[zeros]
    y1 <- bg.cat.1[zeros] - neg.catalog[zeros]
    if (attr(target.catalog, "catalog.type") == "density") {
      # message("adjusting for density")
      y0 <- y0 * 1e6
      y1 <- y1 * 1e6
    }

    # message(paste(y0, collapse = "y"))
    # message(paste(y1, collapse = "z"))
    minus.col <- "purple"
    segments(x0 = bpz, 
             y0 = y0, 
             y1 = y1, 
             col = minus.col,
             lwd = 3,
             lty = 1)

    legend.txt <- c(params$whichnitro, "Background")
    legend.col <- c("grey35", "red")
    if (length(bpz) > 0) {
      legend.txt <- c(legend.txt, paste("Background -", params$whichnitro))
      legend.col <- c(legend.col, minus.col)
    }
    legend("topright", legend = legend.txt, fill = legend.col, cex = 0.8)
  }
}
```

## Generate and read input spectra

These are the spectra from cell lines exposed to cisplatin, carboplatin, or oxaliplatin

### Generate spectra "catalogs" from HepG2 and MCF-10A cisplatin VCF files

```{r, process.cisplatin.vcfs}
if (!file.exists("example.spectra/cis.spectra.Rdata")) {
  vcfs <- list.files("example.vcfs/HepG2_Cis/", full.names = TRUE)
  hepg2.cis <- ICAMS::StrelkaSBSVCFFilesToCatalog(
    files      = vcfs,
    ref.genome = BSgenome.Hsapiens.1000genomes.hs37d5,
    region     = "genome")$catSBS96
  
  vcfs <- list.files("example.vcfs/MCF10A_Cis/", full.names = TRUE)
  mcf10a.cis <- ICAMS::StrelkaSBSVCFFilesToCatalog(
    files      = vcfs,
    ref.genome = BSgenome.Hsapiens.1000genomes.hs37d5,
    region     = "genome")$catSBS96
  
  save(hepg2.cis, mcf10a.cis, file = "example.spectra/cis.spectra.Rdata")
  rm(vcfs)
} else {
  load("example.spectra/cis.spectra.Rdata")
}

```

### Read carboplatin and oxalipltin spectra catalogs

```{r, read.catalogs}
hepg2.car  <- ICAMS::ReadCatalog("example.spectra/HepG2_Car.csv", 
                                 ref.genome = BSgenome.Hsapiens.1000genomes.hs37d5,
                                 region = "genome",
                                 catalog.type = "counts")

hepg2.oxa  <- ICAMS::ReadCatalog("example.spectra/HepG2_Oxa.csv", 
                                 ref.genome = BSgenome.Hsapiens.1000genomes.hs37d5,
                                 region = "genome",
                                 catalog.type = "counts")

mcf10a.car <- ICAMS::ReadCatalog("example.spectra/MCF10A_Car.csv", 
                                 ref.genome = BSgenome.Hsapiens.1000genomes.hs37d5,
                                 region = "genome",
                                 catalog.type = "counts")

```

### Input as count spectra

```{r, plot.input.hepg2.cis, fig.height=7, fig.width=7}
pcat(hepg2.cis)
```

```{r, plot.input.mcf10a.cis, fig.height=10, fig.width=7}
pcat(mcf10a.cis)
```

```{r, plot.input.hepg2.car, fig.height=5, fig.width=7}
pcat(hepg2.car)
```

```{r, plot.input.mcf10a.car, fig.height=7, fig.width=7}
pcat(mcf10a.car)
```

```{r, plot.input.oxa, fig.height=5, fig.width=7}
pcat(hepg2.oxa)
```


`r ifelse(params$verbose, "### Input as density spectra", "")`

```{r input_spectra_as_density, eval=FALSE, echo=params$verbose}
pcat(ICAMS::TransformCatalog(spectra, 
                             target.catalog.type = "density"))
```

## Decompose the background and target signatures

```{r, separate.hepg2}
one.separation <- function(sig.name, sig.list, bg.sig.info, my.seed = 101010) {
  
  set.seed(my.seed, kind = "L'Ecuyer-CMRG")
  
  spectra <- sig.list[[sig.name]]
  
  if (is.null(attr(spectra, "abundance"))) {
    stop("NULL abundance for ", sig.name)
  }
  
  ret <- SeparateSignatureAndSpectra(
    spectra          = spectra,
    bg.sig.info      = bg.sig.info,
    m.opts           = NULL,
    start.b.fraction = 0.5,
    sig.name         = sig.name
  )

  return(ret)
}

hepg2.exposures <- list(hepg2.cis, hepg2.car, hepg2.oxa)
names(hepg2.exposures) <- c("HepG2.cis.sig", "HepG2.car.sig", "HepG2.oxa.sig")
hepg2.separated <- lapply(X           = names(hepg2.exposures),
                          FUN         = one.separation,
                          sig.list    = hepg2.exposures,
                          bg.sig.info = background.info[["HepG2"]])
names(hepg2.separated) <- names(hepg2.exposures)
```

```{r, separate.mcf10a}

mcf10a.exposures <- list(mcf10a.cis, mcf10a.car)
names(mcf10a.exposures) <- c("MCF10A.cis.sig", "MCF10.car.sig")
mcf10a.separated <- lapply(X           = names(mcf10a.exposures),
                          FUN         = one.separation,
                          sig.list    = mcf10a.exposures,
                          bg.sig.info = background.info[["MCF10A"]])
names(mcf10a.separated) <- names(mcf10a.exposures)

```


## Signature based on inferred target spectra plus uncertainty


```{r, ex1}
par(mfrow = c(1, 1))
par(mar = c(7, 5, 7, 2))
par(cex = 0.8)
par(cex.main = 1.4) 
for (test.name in names(hepg2.separated)) {
  PlotSpectraAsSigsWithUncertainty(
    hepg2.separated[[test.name]]$inferred.target.spectra,
    title = test.name)
}


```


### Signature based on subtracted spectra compared to directly inferred signature


```{r, cossim}
cossim <- function(v1, v2) {
  if (!is.null(ncol(v1)))  {
    stopifnot(ncol(v1) == 1)
    v1 <- v1[ , 1]
  }
  if (!is.null(ncol(v2)))  {
    stopifnot(ncol(v2) == 1)
    v2 <- v2[ , 1]
  }
  lsa::cosine(v1, v2)
  
}
```


```{r, sub_inferred_vs_direct_inferred}

sigx <- MeanOfSpectraAsSig(subtracted.target.spectra)$mean.sig
cossim(sigx, inferred.sig)

```


### Uncertainty in inferred signature based on variability of spectra minus inferred background




```{r, dens_var_infer_sig, fig.height = 2.5, warning=FALSE}
d.sub.target.sig.ests <- 
  ICAMS::TransformCatalog(sub.target.sig.ests, 
                          target.catalog.type = "density.signature")

arrow.tops <- apply(d.sub.target.sig.ests, 1, max)
arrow.bottoms <- apply(d.sub.target.sig.ests, 1, min)
min.sub <- min(arrow.bottoms)
bp <- pcat1(
  ICAMS::TransformCatalog(inferred.sig,
                          target.catalog.type = "density.signature"),
  ylim = c(min.sub, max(arrow.tops)  + 0.005))
add.arrows(bp, arrow.tops, arrow.bottoms)

```


### Show combined background and target spectra

```{r, plot.combined.count}
# debug(stacked.pcat)
stacked.pcat(inferred.bg.spectra, subtracted.target.spectra)
```

```{r, plot.combined.density}
stacked.pcat(
  ICAMS::TransformCatalog(inferred.bg.spectra,
                          target.catalog.type = "density"),
  ICAMS::TransformCatalog(subtracted.target.spectra,
                          target.catalog.type = "density"))

```

`r ifelse(params$showrecon, "### Reconstruct spectra", "")`

```{r, get.reconstructed.spectra, eval=params$showrecon, echo=params$showrecon}
reconstructed.target.spectra <- 
  round(inferred.sig %*% matrix(ret$exposures.to.target.sig, nrow = 1))
reconstructed.target.spectra <- 
  ICAMS::as.catalog(
    reconstructed.target.spectra,
    catalog.type   = "counts",
    region         = "genome",
    abundance      = attr(spectra, "abundance"),
    infer.rownames = TRUE)
colnames(reconstructed.target.spectra) <-
  paste0(colnames(spectra), "-mult-recon-spect")
pcat(reconstructed.target.spectra)


total.spectra <- reconstructed.target.spectra + inferred.bg.spectra


lapply(1:ncol(spectra), 
        function(col.index) {
          return(list(
            cossim = cossim(total.spectra[ , col.index], spectra[ , col.index]),
            e.dist = dist(rbind(total.spectra[ , col.index], spectra[ , col.index]), method = "euclidean")))})

```

`r ifelse(params$showrecon, "### Original and reconstructed spectra side-by-side", "")`

```{r, plot.orig.and.recon, eval=params$showrecon, echo=params$showrecon}
InterleaveMatrices <- function(m1, m2) {
  stopifnot(dim(m1) == dim(m2))
  i2 <- 1:ncol(m1) * 2
  i1 <- 1 + (0:(ncol(m1) - 1) * 2)
  out <- matrix(nrow = nrow(m1), ncol = 2 * ncol(m1))
  out[ , i1] <- m1
  out[ , i2] <- m2
  c.names <- character(2 * ncol(m1))
  c.names[i1] <- colnames(m1)
  c.names[i2] <- colnames(m2)
  colnames(out) <- c.names
  rownames(out) <- rownames(m1)
  return(out)
  
}

s.by.s <- InterleaveMatrices(spectra, total.spectra)
s.by.s <- ICAMS::as.catalog(s.by.s, 
                            region = "genome",
                            catalog.type = "counts",
                            abundance = attr(spectra, "abundance"))
```

```{r, plot.replicate1, eval=params$showrecon, echo=params$showrecon}
for (i in 0:(ncol(spectra) - 1)) {
  pcat(s.by.s[ , (2 * i + 1):(2 * i + 2) ])
}
```

## Sanity check: are both the background and the inferred signature needed?

### Is the inferred signature needed?

```{r, sanity_check_1_is_target_needed, eval = TRUE}
library(mSigAct)
m.opts <- mSigAct::DefaultManyOpts()
two.sigs2 <- two.sigs
two.sigs2[ , 1] <- two.sigs2[ , 1] / sum(two.sigs2[ , 1])
mSigAct::SignaturePresenceTest(
  spectra          = spectra,
  sigs             = two.sigs2,
  target.sig.index = 1,
  eval_f           = mSigAct::ObjFnBinomMaxLHNoRoundOK,
  m.opts           = m.opts)

```


### Is the background signature needed?

```{r, sanity_check_2_is_backgrouind_needed, eval = TRUE}
mSigAct::SignaturePresenceTest(
  spectra          = spectra,
  sigs             = two.sigs2,
  target.sig.index = 2,
  eval_f           = mSigAct::ObjFnBinomMaxLHNoRoundOK,
  m.opts           = m.opts)
```